---
title: "Using Sobol with an ODE"
author: "Maxwell Patterson"
date: "2024-05-17"
---

# The Model
Consider the following model of forest growth (where forest size in measured in units of carbon (C))

$$\frac{dC}{dt} = r ∗ C$$ for forests where C is below a threshold canopy closure
$$\frac{dC}{dt} = g ∗ (1 − \frac{C}{K})$$ for forests where carbon is at or above the threshold canopy closure
and $K$ is a carrying capacity in units of carbon

The size of the forest (C), Canopy closure threshold and carrying capacity are all in units of carbon You
could think of the canopy closure threshold as the size of the forest at which growth rates change from
exponential to linear You can think of r, as early exponential growth rate and g as the linear growth rate
once canopy closure has been reached

# Your Task

1. Implement this model in R (as a differential equation)

```{r}
dforestgrowth <- function(Time, C, params, thresh) {
  if (C < thresh) {
    dC_dT <- params$r * C
    return(list(dC_dT))
  } else if (C >= thresh) {
    dC_dT <- params$g * (1 - C / params$K)
    return(list(dC_dT))
  }
}
```


2. Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C, and
using the following parameters:
- canopy closure threshold of 50 kgC
- K = 250 kg C (carrying capacity)
- r= 0.01 (exponential growth rate before before canopy closure)
- g = 2 kg/year (linear growth rate after canopy closure)

```{r}
library(deSolve)
library(tidyverse)
library(sensitivity)
```

```{r}
# Define time, initial forest size, and threshold
time <- seq(1, 300)
Cinitial <- 10
thresh <- 50

# Turn params into a list
parms <- list(K = 250, r = 0.01, g = 2)
result <- ode(y = Cinitial, 
              times = time,
              func = dforestgrowth,
              parms = parms,
              thresh = thresh)

# Turn results into a df
result <- as.data.frame(result)
```


3. Graph the results. Here you are graphing the trajectory with the parameters as given (e.g no uncertainty)

```{r}
ggplot(result, aes(time, `1`)) +
  geom_point() +
  theme_linedraw() +
  labs(x = "Time (Years)", y = "Forest Size (kgC)")
```


4. Run a Sobol global (vary all parameters at the same time) sensitivity analysis that explores how the
estimated maximum forest size (e.g maximum of C 300 years, varies with these parameters
- pre canopy closure growth rate (r)
- post-canopy closure growth rate (g)
- canopy closure threshold and carrying capacity(K)
Assume that parameters are all normally distributed with means as given above and standard deviation of
10% of mean value

```{r}
# Generate sobol parameters, st dev is 10% of mean value
num_samples <- 2000
carrying_capacity <- rnorm(mean = 250, sd = 250 * 0.01, n = num_samples)
pre_closure_rate <- rnorm(mean = 0.01, sd = 0.01 * 0.01, n = num_samples)
post_closure_rate <- rnorm(mean = 2, sd = 2 * 0.01, n = num_samples)
param_set1 <- cbind.data.frame(r = pre_closure_rate, K = carrying_capacity, g = post_closure_rate)

# Do for second set of samples
carrying_capacity <- rnorm(mean = 250, sd = 250 * 0.01, n = num_samples)
pre_closure_rate <- rnorm(mean = 0.01, sd = 0.01 * 0.01, n = num_samples)
post_closure_rate <- rnorm(mean = 2, sd = 2 * 0.01, n = num_samples)
param_set2 <- cbind.data.frame(r = pre_closure_rate, K = carrying_capacity, g = post_closure_rate)

# Fix negative vals
param_set1 <- param_set1 %>% map_df(pmax, 0.0)
param_set2 <- param_set2 %>% map_df(pmax, 0.0)
```


```{r}
# Run sobol model
sobol_result <- sobolSalt(model = NULL, X1 = param_set1, X2 = param_set2, nboot = 300)

# Add column names
colnames(sobol_result$X) <- c("r", "K", "g")

# Get results for 300 years (evaluating every year)
canopy_threshold <- 50
simulation_times <- seq(from = 1, to = 300)
params <- list(r = sobol_result$X[1, "r"],
               K = sobol_result$X[1, "K"],
               g = sobol_result$X[1, "g"])

# Run ode solver
ode_result <- ode(y = Cinitial,
                  times = simulation_times,
                  func = dforestgrowth,
                  parms = params,
                  thresh = canopy_threshold)

ode_result <- as.data.frame(ode_result)

# Rename the column names
colnames(ode_result) <- c("time", "C")
```

```{r}
# Create metrics function
calculate_metrics <- function(result, threshold) {
  max_growth <- max(result$C)
  
  index <- which(result$C > threshold)[1] 
  index <- ifelse(is.na(index), 
                  length(result$C), 
                  index) 
  
  threshold_year <- result$time[index]
  
  return(list(maxgrowth = max_growth, threshyear = threshold_year))
}

# Apply function to first param set
param_wrapper <- function(r, K, g, Cinitial, simtimes, odefunc, metricfunc, threshold) {
  params <- list(r = r, K = K, g = g)
  result <- ode(y = Cinitial, 
                times = simtimes, 
                func = odefunc, 
                parms = params,
                thresh = threshold)
  colnames(result) <- c("time", "C")
  
  # Pull metrics
  metrics <- metricfunc(as.data.frame(result), thresh = 100)
  return(metrics)
}
```

```{r}
all_metrics <- as.data.frame(sobol_result$X) %>% 
  pmap(param_wrapper, 
       thresh = 50,
       Cinitial = Cinitial, 
       simtimes = time,
       odefunc = dforestgrowth,
       metricfunc = calculate_metrics)

# Get results into a df
all_metrics_df <- all_metrics %>% 
  map_dfr(`[`, c("maxgrowth", "threshyear"))
```

```{r}
# Build boxplots
tmp_data <- all_metrics_df %>% 
  pivot_longer(cols = everything(),
               names_to = "metric", 
               values_to = "value")

ggplot(tmp_data, aes(metric, value, color = metric)) + 
  geom_boxplot() + 
  theme_linedraw()
```

```{r}
# Get sobol indices
sobol_maxgrowth <- sensitivity::tell(sobol_result, all_metrics_df$maxgrowth)

# First-order indices (main effect without co-variance)
rownames(sobol_maxgrowth$S) <- c("r", "K", "g")
sobol_maxgrowth$S

# Total sensitivity index - note that this partitions the output variance
rownames(sobol_maxgrowth$T) <- c("r", "K", "g")
sobol_maxgrowth$T

# Get sobol indices for threshold year
sobol_threshyear <- sensitivity::tell(sobol_result, all_metrics_df$threshyear)

# First-order indices (main effect without co-variance)
rownames(sobol_threshyear$S) <- c("r", "K", "g")
sobol_threshyear$S

# Total sensitivity index - note that this partitions the output variance - so values sum to 1
rownames(sobol_threshyear$T) <- c("r", "K", "g")
sobol_threshyear$T
```




5. Graph the results of the sensitivity analysis as a box plot of maximum forest size and record the two
Sobol indices (S and T).

```{r}
# Boxplots
tmp_data <- all_metrics_df %>%
  pivot_longer(cols = everything(),
               names_to = "metric",
               values_to = "value")

ggplot(tmp_data, aes(metric, value, color = metric)) +
  geom_boxplot() +
  theme_linedraw()

# Get sobol indices for max growth
sobol_maxgrowth <- sensitivity::tell(sobol_result, all_metrics_df$maxgrowth)

# First-order indices 
rownames(sobol_maxgrowth$S) <- c("r", "K", "g")
sobol_maxgrowth$S

# Total sensitivity index 
rownames(sobol_maxgrowth$T) <- c("r", "K", "g")
sobol_maxgrowth$T

# Get sobol indices for threshold year
sobol_threshyear <- sensitivity::tell(sobol_result, all_metrics_df$threshyear)

# First-order indices 
rownames(sobol_threshyear$S) <- c("r", "K", "g")
sobol_threshyear$S

# Total sensitivity index, values sum to 1
rownames(sobol_threshyear$T) <- c("r", "K", "g")
sobol_threshyear$T
```

The main effects for the parameters influencing maximum forest growth are 0.329 (r), 0.416 (K), and 0.227 (g). The total sensitivity indices for maximum growth rate are 0.340 (r), 0.406 (K), and 0.243 (g).

The main effects for threshold year for the parameters influencing maximum forest growth are 0.917 (r), 0.0331 (K), and 0.0671 (g). The sensitivity for maximum forest growth are 0.937 (r), 
0.0306 (K), and 0.0750 (g).



6. In 2-3 sentences, discuss what the results of your simulation might mean. (For example think about how what parameters climate change might influence).

Looking at the results of the sensitivity analysis, maximum forest growth is most heavily influenced by the carrying capacity and pre-closure growth rate. The threshold year is primarily determined by the pre-closure growth rate (r). Climate change could impact these parameters since  changes in temperature and precipitation patterns are likely to affect forest carrying capacity and growth rates Understanding and analyzing the sensitivity of forest growth to these parameters can help us predict, manage and understand the potential response of forests to a range of future climate scenarios.


