---
title: "Using Sobol with an ODE"
author: "Maxwell Patterson"
date: 2024-05-17
---

# The Model
Consider the following model of forest growth (where forest size in measured in units of carbon (C))

$$\frac{dC}{dt} = r ∗ C$$ for forests where C is below a threshold canopy closure
$$\frac{dC}{dt} = g ∗ (1 − \frac{C}{K})$$ for forests where carbon is at or above the threshold canopy closure
and $K$ is a carrying capacity in units of carbon

The size of the forest (C), Canopy closure threshold and carrying capacity are all in units of carbon You
could think of the canopy closure threshold as the size of the forest at which growth rates change from
exponential to linear You can think of r, as early exponential growth rate and g as the linear growth rate
once canopy closure has been reached

# Your Task

1. Implement this model in R (as a differential equation)

```{r}
forest_growth <- function(t, C, params) {
  pre_canopy_closure_rate <- params[[1]]
  post_canopy_closure_rate <- params[[2]]
  canopy_closure_threshold <- params[[3]]
  carrying_capacity <- params[[4]]
  
  if (C < canopy_closure_threshold) {
    dC_dt <- pre_canopy_closure_rate * C
  } else {
    dC_dt <- post_canopy_closure_rate * (1 - C/carrying_capacity)
  }
  
  return(list(dC_dt))
}
```


2. Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C, and
using the following parameters:
- canopy closure threshold of 50 kgC
- K = 250 kg C (carrying capacity)
- r= 0.01 (exponential growth rate before before canopy closure)
- g = 2 kg/year (linear growth rate after canopy closure)

```{r}
library(deSolve)

# Set the initial conditions and parameter values
C0 <- 10  # Initial forest size
times <- seq(0, 300, by = 1)  # Time sequence from 0 to 300 years
params <- list(pre_canopy_closure_rate = 0.01,
               post_canopy_closure_rate = 2,
               canopy_threshold = 50,
               carrying_capacity = 250)

# Solve the ODE
solution <- ode(y = C0, times = times, func = forest_growth, parms = params)
```


3. Graph the results. Here you are graphing the trajectory with the parameters as given (e.g no uncertainty)

```{r}
plot(solution[, 1], solution[, 2], type = "l", xlab = "Time (years)", ylab = "Forest Size (kgC)")
```


4. Run a Sobol global (vary all parameters at the same time) sensitivity analysis that explores how the
estimated maximum forest size (e.g maximum of C 300 years, varies with these parameters
- pre canopy closure growth rate (r)
- post-canopy closure growth rate (g)
- canopy closure threshold and carrying capacity(K)
Assume that parameters are all normally distributed with means as given above and standard deviation of
10% of mean value

```{r}
library(sensitivity)
np <- 2000  # Number of parameter samples

# Generate samples for pre-canopy closure rate (r)
r1 <- rnorm(np, mean = 0.01, sd = 0.001)
r2 <- rnorm(np, mean = 0.01, sd = 0.001)

# Generate samples for post-canopy closure rate (g)
g1 <- rnorm(np, mean = 100, sd = 10)
g2 <- rnorm(np, mean = 100, sd = 10)

# Generate samples for canopy closure threshold
canopy_threshold1 <- rnorm(np, mean = 50, sd = 5)
canopy_threshold2 <- rnorm(np, mean = 50, sd = 5)

# Generate samples for carrying capacity (K)
K1 <- rnorm(np, mean = 250, sd = 25)
K2 <- rnorm(np, mean = 250, sd = 25)

# Combine the samples into a data frame
X1 <- data.frame(r = r1, g = g1, canopy_threshold = canopy_threshold1, K = K1)
X2 <- data.frame(r = r2, g = g2, canopy_threshold = canopy_threshold2, K = K2)
```

```{r}
# Define a function to compute the maximum forest size
compute_max_forest_size <- function(params) {
  # Run the ODE solver with the given parameters
  C0 <- 10
  times <- seq(0, 300, by = 1)
  solution <- ode(y = C0, times = times, func = forest_growth, parms = params)
  
  # Return the maximum forest size
  return(max(solution[, 2]))
}

# Run the Sobol analysis
sens_max_forest_size <- sobolSalt(model = NULL, X1, X2, nboot = 300, fun = compute_max_forest_size)

# Extract the maximum forest size values
max_forest_size <- sens_max_forest_size$Y

# Check for infinite or missing values
any(is.infinite(max_forest_size))
any(is.na(max_forest_size))

# Get the minimum and maximum values of your data
min_value <- min(max_forest_size, na.rm = TRUE)
max_value <- max(max_forest_size, na.rm = TRUE)

# Adjust the limits to provide some padding
ylim_lower <- min_value - 0.1 * min_value
ylim_upper <- max_value + 0.1 * max_value

# Create the boxplot with manually set ylim
boxplot(max_forest_size, main = "Maximum Forest Size", ylim = c(ylim_lower, ylim_upper))

# Extract the Sobol indices
S_indices <- sens_max_forest_size$S
T_indices <- sens_max_forest_size$T
```


5. Graph the results of the sensitivity analysis as a box plot of maximum forest size and record the two
Sobol indices (S and T).

```{r}
# Extract the maximum forest size values
max_forest_size <- sens_max_forest_size$Y

# Create a boxplot
boxplot(max_forest_size, main = "Maximum Forest Size")

# Extract the Sobol indices
S_indices <- sens_max_forest_size$S
T_indices <- sens_max_forest_size$T
```


6. In 2-3 sentences, discuss what the results of your simulation might mean. (For example think about
how what parameters climate change might influence).

# Bonus

Also compute Sobol indices for a second metric: forest size after a 100 years


